---
title: "Rachel's Strava App"
author: "Rachel Swallow"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
    toc: true
    toc_depth: 4
    theme: paper 
    df_print: paged
runtime: shiny
---
One can get fancy with allowing the user a lot of flexibility in terms of how they choose the various filters and splits. But you've captured the main idea nicely. 

Best to change the column headings to something more text like ("Split time" rather than "splittime" etc).

I'm not sure if this is an actual run but if so just double check your distance and time calculations because your speeds seem off. I've only looked at a few runs but they're like 50 mins for 10-12k, so not 1-2 min per km.

Not sure why things are slow to recalculate... once you've filtered out the last few runs, which is fast, then the other operations are just averages, so should be fast too. One thing is that st_distance can be slow, so you might want to code the distance calculation yourself.

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height=5,
                      fig.align='center',              tidy.opts=list(width.cutoff=60),
                      tidy=TRUE, 
                     warning = FALSE,echo=FALSE,comment=NA)
library(knitr)
library(tinytex)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)
library(rmarkdown)
library(ggrepel)
library(reshape2)
library(raster)
library(sf)
```
### Plagerism Declaration
I understand the meaning of plagiarism and I declare that all of the work presented here in this report is my own, unless properly acknowledged. 

### Assignment Description 
\marginpar{\bf{[15]}}

```{r loadingdata}
#Load data into a R object list 
directory <- ("./data/data")
specdatafiles<-as.character(list.files(directory))
specdatapaths <- paste(directory, specdatafiles, sep="/")

files<-list()

for(i in 1:length(specdatapaths))
  {
      files[[i]]<- read.csv(specdatapaths[i], header=T, sep=",")
}

# Save an object to a file
saveRDS(files, file = "my_data.rds")
my_data <- readRDS("~/2021/EDA/EDA workspace/EDA_Assignment3/my_data.rds")
```

```{r}
dateRange1<-ymd("2021-03-01")
my_data1<- lapply(my_data, function(x) {
  x$date <- ymd(x$date)
  x$time<-hms(x$time)
  x
})

r<-lapply(my_data1, function(x) filter(x, date==dateRange1[1]))
   #For loop
   ind<-c()
   j<-1
   for(i in 1:length(my_data1)){
      if (length(r[[i]]$date)>2){
        ind[j]<-i
        j<-j+1
      }
   }
   
b<-my_data1[ind]#gives back a list of non-empty dataframes
b<-as.data.frame(b)

```

```{r method1}
geo<- st_as_sf(b,coords = c("lat","lng","elevation"),crs = 4326)
my_crs <- "+proj=utm +zone=34H +datum=WGS84 +units=m +no_defs" #for Beaufort 
geo <-st_transform(geo, crs = my_crs)
dist<-st_coordinates(geo)
dist<-as.data.frame(dist)
dist<-dist%>%mutate(distance=sqrt((X-lag(X))^2+(Y-lag(Y))^2))
dist$distance[1]<-0
dist<-dist%>%mutate(tot_dist=cumsum(distance)/1000)
```

```{r method2}
geo<- st_as_sf(b,coords = c("lat","lng","elevation"),crs = 4326)
my_crs <- "+proj=utm +zone=34H +datum=WGS84 +units=m +no_defs" #for Beaufort 
geo <-st_transform(geo, crs = my_crs)

dist1<-c()
for(i in 1:(length(geo$geometry)-1)){
  dist1[i]<-st_distance(geo$geometry[i],geo$geometry[i+1],'meter')
}
#dist2<-st_distance(geo,by)
dist2<-st_distance(geo$geometry,by_element = TRUE)
dist1<-c(0,dist)
geo$dist<-dist1
#st_coordinate then dataframe, lag ()
geo<-geo%>%mutate(tot_dist=cumsum(dist)/1000) #in meters 
tot_distance<-max(geo_test$tot_dist)

km<-seq(from=1,to=round(tot_distance),by=1)#initilize coloumn one of dataframe = the kms 

splittime<-c()
 for(i in 1:round(tot_distance)){
      new<-geo%>%filter(tot_dist<i & tot_dist>=(i-1)) #filter out the data between the two kms relevent = 1km split 
      splittime[i]<-round((new$time[length(new$time)]-new$time[1])/60)
 }
    splits_time<-as.data.frame(cbind(km,splittime))
    splits_time
    
    
elev<-max(b$elevation)-min(b$elevation) #elevation in meters
```

```{r}
shinyApp(
  # Define UI for application
ui <- fluidPage(
  
  navbarPage("Rachel's Strava", theme = shinytheme("united"), #Add a navigation panel with a nice theme 
             tabPanel("Activity Data", fluid = TRUE, icon = icon("running"), #Add icons for nav sections 
  #PAGE 1 
  sidebarLayout( #define sideoanel layout 
  sidebarPanel(
    
    titlePanel("Run Statistics"), #Give a title to the side bar and Page 
    
    
                    
    #User can select date of the run they want to plot                
    dateInput('date',
              label = 'Pick a run date: ', #guide the user to pick a date of the run they want to analyze
              value = "2021-03-01", #Start date is set as the last run 
              min="2019-02-27",#Minimum date of run (002.csv)
              max="2021-03-01" #Maximum date of run (176.csv)
    )
    ),
  mainPanel( #in the main panel of the first page plot
    #Draw Map 
    leafletOutput("map"), #UI Map
    #Summary Statistics 
    textOutput("act_t"),
    textOutput("dist"),
    textOutput("elv"),
    textOutput("pace")

    )
)),

tabPanel("Activity Splits", fluid = TRUE, icon = icon("chart-bar"), #Label the second nav page 
         titlePanel("Activity Breakdown "), 
           mainPanel(
             #Display Table of Data Metrics
             dataTableOutput("kmtable")#interactive data table for splits
           )
        )
)
),

# Define server logic required to draw a histogram
server <- function(input, output, session) {
  
  dataset <- reactive({
    r<-lapply(my_data, function(x) filter(x, date==input$date)) #Select the dataframe in the list with the matching date 
    #Filter out all the dataframes in the list with number of rows smaller than 2. Remaining list will only contain the dataframe of the run. 
    ind<-c()
    j<-1
    for(i in 1:length(my_data)){
       if (length(r[[i]]$date)>2){

         ind[j]<-i
         j<-j+1

       }
    }
     as.data.frame(my_data[ind])#gives back a list of non-empty dataframes
   })
  
  #Work with geo data 
  geodata<-reactive({
    #m<-st_as_sf(dataset(),coords = c("lat","lng","elevation"),crs = 4326)
    m<-st_as_sf(dataset(),coords = c("lat","lng"),crs = 4326) #Projection for distance calculations 
    my_crs <- "+proj=utm +zone=34S +datum=WGS84 +units=m +no_defs" #for Beaufort 
    sights<-st_transform(m, crs = my_crs) #transform 
    
    
    dist<-c()
    for(i in 1:(length(sights$geometry)-1)){
      dist[i]<-st_distance(sights$geometry[i],sights$geometry[i+1])#Calculate the distance between points 
    }
    
    dist1<-c(0,dist)#replace index 1 with a zero 
    sights$dist<-dist1 #Add the coloumn to sights 
    
    sights_test<-sights%>%mutate(tot_dist=cumsum(dist)/1000) #cumulative sum in km
    sights_test #return dataframe 
  
  })

  #Work out averages 
  tot_distance<-reactive({
    round(geodata()$tot_dist[length(geodata()$tot_dist)],2)#get last entry - total distance 
  })
  
  acttime<- reactive({
    round((dataset()$time@.Data[length(dataset()$time@.Data)]-dataset()$time@.Data[1])/60,2) #Difference in seconds between first point and last point divided by 60 = minutes 
  })
  
  elev<-reactive({
    max(dataset()$elevation)-min(dataset()$elevation)#Change in elevation 
  })
  
  splits<- reactive({
    km<-seq(from=1,to=round(tot_distance()),by=1)#initilize coloumn one of dataframe = the kms 
    splittime<-c()#empty list 
    for(i in 1:round(tot_distance())){ #for kms of run 
      new<-geodata()%>%filter(tot_dist<i & tot_dist>=(i-1)) #filter out the data between the two kms relevent = 1km split 
      splittime[i]<-round((new$time@.Data[length(new$time@.Data)]-new$time@.Data[1])/60,2)#time differences between the two datapoints representing the km difference 
       
    }
    df<-as.data.frame(cbind(km,splittime))#return a dataframe 
    df
  })
  
  
# PAGE 1__________________________________________________________________________________________________________

  #Renders Text
  output$act_t <- renderText({
    paste("Total Activity Time:" ,acttime(), "mins")
  })
  
  output$dist <- renderText({
    paste("Total Distance: ",tot_distance(), "km")
  })
  
  output$elv <- renderText({
    paste("Total Elevation Gain: ",elev(), "m")
  })
  
  output$pace <- renderText({
    paste("Average Pace :" ,round(acttime()/tot_distance(),2), "min /km" )
  })
  
  #Output Map 
 output$map <- renderLeaflet({
   leaflet() %>% addTiles() %>%addPolylines(lng=dataset()$lng, lat=dataset()$lat,col="#FC4C1A",popup="Running Route")%>%addMiniMap(position = "bottomleft")#says its not subsetable
 })#Plotted in Strava colors 
 
#PAGE 2_________________________________________________________________________________________________________
 
 output$kmtable <- renderDataTable({ #Plot a reactive table 
   splits()
 },
 options = list(  columns = list(
   list(title = 'Kilometer Split'),
   list(title = 'Split time (min)')
 ))
 )
 
},

   options = list(height = 500)
)

```

