---
title: "Rachel's Strava App"
author: "Rachel Swallow"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: html_document
    toc: true
    toc_depth: 4
    theme: paper 
    df_print: paged
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height=5,
                      fig.align='center',              tidy.opts=list(width.cutoff=60),
                      tidy=TRUE, 
                     warning = FALSE,echo=FALSE,comment=NA)
library(knitr)
library(tinytex)
library(ggplot2)
library(dplyr)
library(tidyverse)
library(lubridate)
library(rmarkdown)
library(ggrepel)
library(reshape2)
library(raster)
```
### Plagerism Declaration
I understand the meaning of plagiarism and I declare that all of the work presented here in this report is my own, unless properly acknowledged. 

### Assignment Description 
\marginpar{\bf{[15]}}

```{r loadingdata}
directory <- ("./data/data")
specdatafiles<-as.character(list.files(directory))
specdatapaths <- paste(directory, specdatafiles, sep="/")

files<-list()

for(i in 1:length(specdatapaths))
  {
      files[[i]]<- read.csv(specdatapaths[i], header=T, sep=",")
}

# Save an object to a file
saveRDS(files, file = "my_data.rds")

```

```{r}
dateRange1<-c(ymd("2021-03-01"),ymd("2021-02-11"))
my_data<- lapply(my_data, function(x) {
  x$date <- ymd(x$date)
  x$time<-hms(x$time)
  x #why this ?
})
r<-lapply(my_data, function(x) filter(x, between(date,dateRange1[1],dateRange1[2]) ))
r<-lapply(my_data, function(x) filter(x, date==dateRange1[1] ))
   #For loop
   ind<-c()
   j<-1
   for(i in 1:length(my_data)){
      if (length(r[[i]]$date)>2){
        
        ind[j]<-i
        j<-j+1
        
      }
   }
   
b<-my_data[ind]#gives back a list of non-empty dataframes
b<-as.data.frame(b)
m <- st_as_sf(b,coords = c("lat","lng","elevation"),crs = 4326)
#m %>% st_crs()
## Coordinate Reference System:
##   EPSG: 4326 
##   proj4string: "+proj=longlat +datum=WGS84 +no_defs"
# reprojecting
my_crs <- "+proj=utm +zone=34H +datum=WGS84 +units=m +no_defs" #for Beaufort 
sights <-  st_transform(m, crs = my_crs)

dist<-c()
for(i in 1:(length(sights$geometry)-1)){
  dist[i]<-st_distance(sights$geometry[i],sights$geometry[i+1])#10m from 
}

dist1<-c(0,dist)
sights$dist<-dist1

sights_test<-sights%>%mutate(tot_dist<-cumsum(dist)/1000) #in meters 

```


Ideas 
get_distance(): returns the total track distance, in meters
get_start_time(): returns a Javascript Date object representing the starting time
get_end_time(): returns a Javascript Date object representing when the last point was recorded
get_moving_time(): returns the moving time, in milliseconds
get_total_time(): returns the total track time, in milliseconds
get_moving_pace(): returns the average moving pace in milliseconds per km
get_moving_speed(): returns the average moving speed in km per hour
get_total_speed(): returns the average total speed in km per hour
get_elevation_min(): returns the lowest elevation, in meters
get_elevation_max(): returns the highest elevation, in meters
get_elevation_gain(): returns the cumulative elevation gain, in meters
get_elevation_loss(): returns the cumulative elevation loss, in meters